### 2018-04-24 source/paper.tex
edited 3c295 section
proofread intro

\pg
The problem of calibration is what concerns us in this paper. It consists of estimating and correcting for instrumental errors (which includes effects such as antenna pointing errors, but also the phase-delays caused by ionospheric activity, troposphere, etc). 
 {Calibration consists of solving for gain estimates, where a gain models the relationship between the electromagnetic field of an astrophysical source and the voltage that an antenna measures for this source. Because measurements are noisy, calibration often involves some fine-tuning of solution intervals, to ensure that the solutions are well-constrained while the solution intervals stay as small as signal-to-noise allows.} The calibration inverse problem involves three competing statistical effects: thermal noise in the measurements, true gain variability, and {sky} model incompleteness. If gain solutions {are sought individually for each measurement}, then calibration estimates will be dominated by thermal noise, and will not adequately describe the actual gains. Similarly, if a single gain estimate is fitted to too many measurements, the intrinsic gain variability will be ``averaged out''; for example, a choice of time and frequency interval that is too large will cause the solver to estimate a constant gain while the underlying function varies quickly, thereby missing much of the gain structure. This will introduce error which will be correlated in time and frequency. This occurs, for example, when solving for ionospheric phase delays: in the most extreme case, where the solution interval is significantly larger than the scale of ionospheric fluctuations, its varying phase can average out to zero over the interval in time and frequency. Finally, if the model being fitted is incomplete, unmodeled physical flux will likely be absorbed unpredictably into both the gain solutions and the residual visibilities: this absorption of physical flux into gain solutions is known as source suppression \citep[see][and references therein]{2014MNRAS.439.4030G,2013MNRAS.435..597K}.

@@

\pg
The key finding of the present paper is a fundamental relationship between the covariance matrices of residual visibilities and the map of the covariance in the image-plane: the ``Cov-Cov relationship'' between visibility covariance to image-plane covariance. We show that the pixel statistics in the image-plane are determined by a ``noise-PSF'', convolved with each source in the sky (modeled or not). This noise-PSF is the product of the Fourier transforms of the gain covariance matrix with each cell mapped not from $uv$ space to $lm$ coordinates but rather between their respective covariance spaces - from a new {differential Fourier plane (henceforth ``$\left(\deltu\deltv\right)$''-plane) to the image-plane covariance space $\deltl\deltm$. This image-plane covariance space describes} the variance in each pixel and the covariance between pixels\footnote{{The noise-PSF also relates $\delta w$ to $\delta n$, as shown in the matrix formalism, but this is not explicitly referenced in the text since visibility space is usually referred to as ``the UV-plane'' in literature, rather than ``the UVW-space''.}}. It describes the expected calibration artefacts and thermal noise around each source, does not vary as a function of direction, and is convolved with each source in the field to yield the final error map. Because all unwanted (in our case, unphysical) signal can be thought of as noise, we will refer to {the pixel variance map} as the ``noise-map''.

@@

\pg
The main result of this paper consists of describing a new adaptive, quality-based weighting scheme based on this insight. Using the Cov-Cov relationship, we can {create a new} weighting scheme by estimating the residual visibility covariance matrix in a given observation. By weighting {visibilities} so as to change their covariance matrix, one can change the shape of the noise-PSF and thus improve {the final image}: this manifests as either decreased noise or decreased calibration artefacts. Note that this weighting is applied after calibration, but before image deconvolution: applying it will therefore not only improve the residual noise in the image, and thus the sensitivity achievable with a given pipeline, but will also improve deconvolution by minimising calibration artefacts in the field: it should thus effectively remove spurious, unphysical emission from final data products. Estimating the covariance matrix is the main difficulty of our framework: we do not know the underlying covariance matrix, and the conditioning of our estimation thereof is limited by the number of measurements within each {solution interval}. As such, we have no guarantee that our estimate of the corrected visibility covariance matrix is accurate. This problem can be alleviated, for example {by estimating the covariance matrix for the antenna gains themselves}, and use {it} to build the visibility covariance matrix: this effectively improves conditioning (cf Sec. \ref{section.algorithm}).

\pg
This paper is split into four main sections. In Section \ref{sec.formalism}, we derive the Cov-Cov relationship. With its newfound insights, we propose quality-based weighting schemes with which to improve radio interferometric images in Section \ref{sec.DynamicRange}. We follow in Section \ref{section.algorithm} by showing how to estimate, from real data, the covariance matrix from which the quality-based weights are derived. Our approach seems to give good results. Finally, we close the paper on a discussion of the applicability and limitations of the quality-based weighting scheme.

% !TeX spellcheck = en_US

\begin{table}[t!]
\begin{tabular}{c l}
\hline
&\\
Scalars & \\
\cline{1-1}
&\\
$n_{\mathrm{pix}}$ & Total number of pixels in image-plane \\ 
                   &  {\& number of cells in $uv$-grid}\\
$n_{\mathrm{ant}}$ & Total number of antennas in the array\\
$n_b$     & Number of visibilities\\
$b$       & Index for a single visibility.\\
& Equivalent to $\left(pq,t\nu\right)$\\
$\tnuIndeix$ & Equivalent to $\left(t,\nu\right)$\\
%$\conj{z}$  & Complex conjugate of $z$\\
&\\
Vectors & \\
\cline{1-1}
&\\
$\muely$  & Residual image vector, size $n_{\mathrm{pix}}$\\
$\bm{\epsilon}$       & {Vector of $\epsilon$, size $n_{\mathrm{pix}}$}\\ 
$\bm{\widetilde{\gamma}}$ & Contains  gain products, size $n_b$. See Eq. \ref{eq.residuals.definition}\\
$\mathbf{1}$ & Vector containing 1 in every cell, size $n_b$\\
$\bm{\deltu}_{bb'}$ & Vector of coordinates in differential Fourier\\ & plane, of length $3$.\\
$\bm{l}_d$          & Vector of sky coordinates, of length $3$.\\
&\\
Matrices&\\
\cline{1-1}
&\\
$\matVt$       & Visibility seen by a baseline $pq$ at time and \\
& frequency $t,\nu$. Size $2\times 2$.\\
$\Kmat{p,t\nu}^d$& Fourier kernel for direction $d$, antenna $p$ and\\
                 &  one $\left(t,\nu\right)$ pair. Size $2 \times 2$.\\
$\matJ{p}$     & Jones matrix for antenna $p$ for one $\left(t,\nu\right)$ pair.\\
& Contains the gains  information. Size $2 \times 2$\\
$\Bmat{}$        & Sky brightness distribution matrix, of size $2 \times 2$.\\
$\bm{N}$       & Noise matrix, of size $2\times 2$. Contains a single \\
& realisation $n$ of the thermal noise in each cell.\\
$\muelF$       & Fourier transform matrix, of size $n_{\mathrm{pix}}\times n_{\mathrm{pix}}$.\\
$\muelS_b$     & Baseline selection matrix, which picks out 1\\
&  visibility out of the full set. Size $n_{\mathrm{pix}} \times n_b$\\
$\muelC_b$     & $n_{\mathrm{pix}} \times n_{\mathrm{pix}}$ convolution kernel that defines\\
               &  the PSF.\\
$\muelF_{bb'}$ & Convolution matrix mapping one $\deltu\deltv$ to $\deltl\deltm$.\\
& The set of all $\muelF_{bb'}$ determines the noise-PSF.\\
               & Size $n_{\mathrm{pix}} \times n_{\mathrm{pix}}$.\\
&\\
\hline
\end{tabular}
\caption{Table recapitulating the meaning and dimensions of vectors and matrices used {in Sec. \ref{sec.dudv}}. Only scalars which give matrix dimensions or indices are given here.}
\label{tab.variables}
\end{table}

@@

\begin{figure*}[!t]
\centering
\includegraphics[width=\textwidth]{images/psfIM-dudv-labelled.png}
\caption{\label{plot.psf.uv.dudv} Fig. \ref{plot.psf.uv.dudv}a shows the PSF image of a simulated 1Jy source at phase centre. {Colourbar units are in Jansky}.  Fig. \ref{plot.psf.uv.dudv}b shows the associated UV track, and Fig. \ref{plot.psf.uv.dudv}c the corresponding $\left(\deltu,\deltv\right)$ tracks. Note that the $\deltu\deltv$ plane does not have a homogeneous point density, but is denser near its origin: here, this is shown by plotting only 1 random point in 10000.}
\end{figure*}

@@

\pg
This $\deltu\deltv$ space warrants further discussion: Fig. \ref{plot.psf.uv.dudv} shows, for a given $uv$-track (Fig. \ref{plot.psf.uv.dudv}b), both the corresponding $\deltu\deltv$ domain (Fig. \ref{plot.psf.uv.dudv}c) and point-spread function (Fig. \ref{plot.psf.uv.dudv}a). {The symmetric, negative $uv$-track is treated as a separate track, and thus ignored. This means that we do not fully constrain the noise-PSF (since the covariance matrix of the symmetric track is simply the Hermitian of the first), but we do not seek to constrain it in this section, but rather to show that our results hold.}. We can see that the $\delta u \delta v$-tracks are symmetrical about the origin. The $\delta u \delta v$ space corresponding to a given $uv$-track can thus be most concisely described as a ``filled $uv$-track", with its boundaries defined by the ends of the $uv$-track. The set of $\muelF_{bb'}$, each of which maps one value of the covariance matrix to a fringe in the image-plane, would then characterise a PSF equivalent for the noise distribution, which we refer to as the noise-PSF. In our formalism, the only source of statistical effects in the field are calibration errors and thermal noise. The average variance in all pixels will be given by the diagonal of the covariance matrix and the thermal noise, provided that it truly follows a normal distribution. The only effects which will cause the variance in the image-plane to vary from one pixel to the next are those mapped onto the covariance fringes, i.e. such position-dependent variance fluctuations will be caused by correlated gain errors, which are spurious signal introduced by erroneous gain estimates. Assuming all sources in the field are point-like and distant, then these variance fluctuations will follow a specific distribution, convolved to every source in the field.
{Since the variance fluctuations act as tracers for calibration artefacts, artefacts in the image can be understood as \emph{one realisation} of the variance map, which is characterised by an average level determined by the variance in the gains and thermal noise, and a noise-PSF convolved with the sky brightness distribution}. The actual artefacts in the image will still be noisy, as a realisation of the true variance map. For the same reason, in the absence of correlated gain errors, $[\Cov{\matGains}]_{b \ne b'}$ are all zero and $\Cov{\matGains}$ is a diagonal matrix. {We then recover a ``flat" noise-map: the variance will be the same in all pixels, as the noise-PSF is absent}. In the ideal case, were we to recover the true value of the gains for all times and frequencies, this becomes pure thermal noise.
\subsection{Noise Map Simulations}\label{sec.simulations}

\begin{figure}[t!]
\centering
\includegraphics[width=0.5\textwidth]{images/CovarianceMatrix.png}
\caption{\label{fig.covmat} Example of {a non-stationary covariance matrix}, which can be used to simulate $\Cov{\matGains}$.{The colourbar units are Jy$^2$}. The correlation scale $\sigma_\tnuIndeix$ is 40 cells, and the variability period is 500 cells. The matrix is made positive semi-definite (and therefore a covariance matrix) through SVD decomposition. The maximum size of the ``bubbles" is determined by $\sigma_\tnuIndeix$.}
\end{figure}

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=\textwidth]{images/cross_sections1.png}
\caption{\label{plot.cross-sections} The three lines in each figure correspond to three horizontal cross-sections from images in Fig. \ref{plot.noisepsfs}. {The units on the y-axis are dimensionless [$Jy^2/Jy^2$].} $\sigma_\tnuIndeix$ is the maximum characteristic error correlation length. In decreasing intensity, they correspond to  $m=0''$, $m=4''$, and $m=8''$. The dashed lines correspond to the variance measured with 2000 realisations for each pixel, while the solid line corresponds to the predicted value at that pixel. There are 31 pixels. We do not show cross-sections for negative $m$ due to image symmetry about the origin.}
\end{figure*}

\begin{figure}[t!]
\centering
\includegraphics[width=0.5\textwidth]{images/noisepsf-ctime400.png} %{../../../SkyCoverageMap.png} %
\caption{\label{plot.noisepsfs} Simulated noise-maps, compared with theoretical prediction. The pixel values are normalised by the average value of the covariance matrix: the units of the colourbar are thus dimensionless ($Jy^2/Jy^2$). These are on the same angular scale as the source shown in Fig. \ref{plot.psf.uv.dudv}.}
\end{figure}

@@

@@

\begin{figure}[t!]
\centering
\includegraphics[width=0.5\textwidth]{images/Ctime6400-noisePSFandDirty-uncorr.png}
\caption{\label{imag.simu-3sources.noisemap} {Noise-map of sky with correlated gain errors and three point sources. The colourbars of (a), (b) and (c) have dimensionless units, while that of (d) is in Jansky. Note the presence of structure in the residuals (c): these show the limits of our hypothesis that sources are spatially incoherent.}}
\end{figure}

@@

@@

\pg
We find that one $\bm{w}$ which satisfies the above is:
\begin{align}
\bm{w} &= \Cov{\matGains}^{-1} \mathbf{1}\label{eq.fullweights.analytical}
\end{align}
where $\mathbf{1}$ is a vector of ones. These weights depend only on calibration quality: badly-calibrated cells will include spurious time-correlated signal introduced by trying to fit the noise $n$ on visibilities. Down-weighting these cells helps suppress artefacts in the field, at the cost of far-field sensitivity. This weighting scheme is thus only a function of the relative quality of calibration at different times, boosting better-calibrated visibilities and suppressing poorly-calibrated visibilities. For the remainder of this paper, we will refer to these weights as \emph{artefact-optimal} weighting.

@@

\pg
We define the residual matrix as:
\begin{align}
\Rmat{\tnuIndeix} =& \sum_d s_d \Kmat{d,\tnuIndeix} \left( \ghatvect_\tnuIndeix\ghatvect_\tnuIndeix^H - \gvect_\tnuIndeix\gvect_\tnuIndeix^H \right) \Kmat{d,\tnuIndeix}^H + \epsilon \label{eq.residual.matrix}
\end{align}
where we {explicitly place ourselves in the limits of our formalism, i.e.} that we do not have direction-dependent gains. We now see that at the core of Eq. \ref{eq.residual.matrix} lies $\displaystyle\VarmatTau$, where $\sum_{\tnuIndeix\in\cellindex} \VarmatTau = \VarmatC$.
The K-matrix is defined as follows:
\begin{align}\label{kmat.def}
\Kmat{d,\tnuIndeix} =&
\begin{pmatrix}
k_{p,\tnuIndeix}^{d}   &                       &   0     &\\
                       & k_{q,\tnuIndeix}^{d}  &        &\\ 
         0              &                       & \ddots 
\end{pmatrix}
\end{align}
Since the residual matrix depends on the gains, we define the residual visibility vectors as:
\begin{align}
\Rmat{{\cellindex}} =&   \begin{pmatrix} \vdots \\ \Rmat{\tnuIndeix\in \cellindex} \\ \vdots \end{pmatrix} \\
\bm{R} =&   \begin{pmatrix} \hdots \quad \Rmat{c} \quad \hdots \end{pmatrix}
\end{align}

@@

\begin{figure*}[t!]
\centering
\begin{subfigure}{.43\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/paperfig-goodcal.png}}
\caption{\label{image.3c295.goodcal} Well-calibrated, unweighted restored image of the sky near the centre of the Extended Groth Strip. Used for comparison with the other images. {Units of colourbar are Janskys}. This image was made with data calibrated following best practice (solution intervals of 8 seconds, half the bandwidth). rms=5.87mJy/beam}
\end{subfigure}
\hfill
\begin{subfigure}{.43\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/paperfig-noweights.png}}
\caption{\label{image.3c295.nocorr} Poorly-calibrated, unweighted restored image of the sky near the centre of the Extended Groth Strip. {Units of colourbar are Janskys}. This image was made with the same data as for Fig. \ref{image.3c295.goodcal}, but averaged in time and calibrated using larger gain solution intervals: 2 minutes and half the bandwidth. rms=86.4mJy/beam}
\end{subfigure}
\hfill
\begin{subfigure}{.43\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/paperfig-cyrilweights.png}}
\caption{\label{image.3c295.lightcorr} Image made using the same imaging parameters and corrected visibilities as Fig. \ref{image.3c295.nocorr}, using sensitivity-optimal weighting. {Units of colourbar are Janskys}. rms=9.69mJy/beam}
\end{subfigure}
\hfill
\begin{subfigure}{.43\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/paperfig-fullweights.png}}
\caption{\label{image.3c295.fullcorr} Image made using the same imaging parameters and corrected visibilities as Fig. \ref{image.3c295.nocorr}, using artefact-optimal weighting. {Units of colourbar are Janskys}. rms=15.8mJy/beam}
\end{subfigure}
\caption{\label{image.fourRealImages}Restored images of the centre of the Extended Groth Strip, as seen with an 8-hour observation using the full LOFAR array. Fig. \ref{image.3c295.goodcal} shows an image of the field made with good calibration intervals. Fig. \ref{image.3c295.nocorr} shows an image of the field made with poor calibration intervals. Fig. \ref{image.3c295.lightcorr} shows image made with the same visibilities and imaging parameters, but with the application of the sensitivity-optimal weighting scheme. Fig. \ref{image.3c295.fullcorr}, similarly, differs from Fig \ref{image.3c295.lightcorr} only in that artefact-optimal weights, rather than sensitivity-optimal weights, were used. {The histograms of pixel values in each image have 1000 flux bins ranging from -0.16 Jy to 0.16 Jy. Their ordinates are in log scale.} Pixel size is 1.5'' in all images.}

\end{figure*}

@@

\begin{figure}[!t]
    \centering
        \centering
        \includegraphics[width=\linewidth]{images/ctime001-NoisePsfCrossections-backup.png} 
        \caption{The sensitivity-optimal (green) and artefact-optimal (red) weights both give improvements over the unweighted noise-map (blue).} \label{fig.simu.weightapplied}
\end{figure}

@@

@@

\begin{figure*}
\begin{subfigure}{.43\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/paperfig-goodcal.png}}
\caption{\label{image.3c295.goodcal-2} Well-calibrated, unweighted restored image of the sky near the centre of the Extended Groth Strip. Used for comparison with the other images. Calibration solution intervals used were 8 seconds, half the bandwidth. rms=5.87mJy/beam}
\end{subfigure}
\hfill
\begin{subfigure}{.43\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/paperfig-antweights.png}}
\caption{\label{image.3c295.antcorr} Image made using the same imaging parameters and corrected visibilities as Fig. \ref{image.3c295.nocorr}, with the application of antenna-based, sensitivity-optimal weighting. Solution interval of 2 minutes, half the bandwidth. rms=6.69mJy/beam}
\end{subfigure}
\caption{\label{image.twoRealImages}Comparison between the well-calibrated image (i.e. the same image as Fig \ref*{image.3c295.goodcal}) and antenna-based sensitivity-optimal weights.  {Units of both colourbars are Jansky}. We see that we recover a very similar image, despite the fact that the data used for the weighted image are averaged by a factor of 8 compared to those used for the unweighted image.}
\end{figure*}

@@

>>>>

\pg
The problem of calibration is what concerns us in this section. It consists of estimating and correcting for instrumental errors (which includes effects such as antenna pointing errors, but also the phase-delays caused by ionospheric activity, troposphere, etc). 
 {Calibration consists of solving for gain estimates, where a gain models the relationship between the electromagnetic field of an astrophysical source and the voltage that an antenna measures for this source. Because measurements are noisy, calibration often involves some fine-tuning of solution intervals, to ensure that the solutions are well-constrained while the solution intervals stay as small as signal-to-noise allows.} The calibration inverse problem involves three competing statistical effects: thermal noise in the measurements, true gain variability, and {sky} model incompleteness. If gain solutions {are sought individually for each measurement}, then calibration estimates will be dominated by thermal noise, and will not adequately describe the actual gains. Similarly, if a single gain estimate is fitted to too many measurements, the intrinsic gain variability will be ``averaged out''; for example, a choice of time and frequency interval that is too large will cause the solver to estimate a constant gain while the underlying function varies quickly, thereby missing much of the gain structure. This will introduce error which will be correlated in time and frequency. This occurs, for example, when solving for ionospheric phase delays: in the most extreme case, where the solution interval is significantly larger than the scale of ionospheric fluctuations, its varying phase can average out to zero over the interval in time and frequency. Finally, if the model being fitted is incomplete, unmodeled physical flux will likely be absorbed unpredictably into both the gain solutions and the residual visibilities: this absorption of physical flux into gain solutions is known as source suppression \citep[see][and references therein]{2014MNRAS.439.4030G,2013MNRAS.435..597K}.

@@

\pg
The key finding of our algorithmic work is a fundamental relationship between the covariance matrices of residual visibilities and the map of the covariance in the image-plane: the ``Cov-Cov relationship'' between visibility covariance to image-plane covariance. We show that the pixel statistics in the image-plane are determined by a ``noise-PSF'', convolved with each source in the sky (modeled or not). This noise-PSF is the product of the Fourier transforms of the gain covariance matrix with each cell mapped not from $uv$ space to $lm$ coordinates but rather between their respective covariance spaces - from a new {differential Fourier plane (henceforth ``$\left(\deltu\deltv\right)$''-plane) to the image-plane covariance space $\deltl\deltm$. This image-plane covariance space describes} the variance in each pixel and the covariance between pixels\footnote{{The noise-PSF also relates $\delta w$ to $\delta n$, as shown in the matrix formalism, but this is not explicitly referenced in the text since visibility space is usually referred to as ``the UV-plane'' in literature, rather than ``the UVW-space''.}}. It describes the expected calibration artefacts and thermal noise around each source, does not vary as a function of direction, and is convolved with each source in the field to yield the final error map. Because all unwanted (in our case, unphysical) signal can be thought of as noise, we will refer to {the pixel variance map} as the ``noise-map''.

@@

\pg
The main result of this this section is a new adaptive, quality-based weighting scheme based on this insight. Using the Cov-Cov relationship, we can {create a new} weighting scheme by estimating the residual visibility covariance matrix in a given observation. By weighting {visibilities} so as to change their covariance matrix, one can change the shape of the noise-PSF and thus improve {the final image}: this manifests as either decreased noise or decreased calibration artefacts. Note that this weighting is applied after calibration, but before image deconvolution: applying it will therefore not only improve the residual noise in the image, and thus the sensitivity achievable with a given pipeline, but will also improve deconvolution by minimising calibration artefacts in the field: it should thus effectively remove spurious, unphysical emission from final data products. Estimating the covariance matrix is the main difficulty of our framework: we do not know the underlying covariance matrix, and the conditioning of our estimation thereof is limited by the number of measurements within each {solution interval}. As such, we have no guarantee that our estimate of the corrected visibility covariance matrix is accurate. This problem can be alleviated, for example {by estimating the covariance matrix for the antenna gains themselves}, and use {it} to build the visibility covariance matrix: this effectively improves conditioning (cf Sec. \ref{section.algorithm}).

\pg
The remainder of this chapter is split into four main sections. In Section \ref{sec.formalism}, we derive the Cov-Cov relationship. With its newfound insights, we propose quality-based weighting schemes with which to improve radio interferometric images in Section \ref{sec.DynamicRange}. We follow in Section \ref{section.algorithm} by showing how to estimate, from real data, the covariance matrix from which the quality-based weights are derived. Our approach seems to give good results. Finally, we close the chapter on a discussion of the applicability and limitations of the quality-based weighting scheme.

% h!eX spellcheck = en_US

\begin{table}[t!]
\begin{tabular}{c l}
\hline
&\\
Scalars & \\
\cline{1-1}
&\\
$n_{\mathrm{pix}}$ & Total number of pixels in image-plane \& number of cells in $uv$-grid\\
$n_{\mathrm{ant}}$ & Total number of antennas in the array\\
$n_b$     & Number of visibilities\\
$b$       & Index for a single visibility.\\
& Equivalent to $\left(pq,t\nu\right)$\\
$\tnuIndeix$ & Equivalent to $\left(t,\nu\right)$\\
%$\conj{z}$  & Complex conjugate of $z$\\
&\\
Vectors & \\
\cline{1-1}
&\\
$\muely$  & Residual image vector, size $n_{\mathrm{pix}}$\\
$\bm{\epsilon}$       & {Vector of $\epsilon$, size $n_{\mathrm{pix}}$}\\ 
$\bm{\widetilde{\gamma}}$ & Contains  gain products, size $n_b$. See Eq. \ref{eq.residuals.definition}\\
$\mathbf{1}$ & Vector containing 1 in every cell, size $n_b$\\
$\bm{\deltu}_{bb'}$ & Vector of coordinates in differential Fourier\\ & plane, of length $3$.\\
$\bm{l}_d$          & Vector of sky coordinates, of length $3$.\\
&\\
Matrices&\\
\cline{1-1}
&\\
$\matVt$       & Visibility seen by a baseline $pq$ at time and frequency $t,\nu$. Size $2\times 2$.\\
$\Kmat{p,t\nu}^d$& Fourier kernel for direction $d$, antenna $p$ and one $\left(t,\nu\right)$ pair. Size $2 \times 2$.\\
$\matJ{p}$     & Jones matrix for antenna $p$ for one $\left(t,\nu\right)$ pair.\\
& Contains the gains  information. Size $2 \times 2$\\
$\Bmat{}$        & Sky brightness distribution matrix, of size $2 \times 2$.\\
$\bm{N}$       & Noise matrix, of size $2\times 2$. Contains a single \\
& realisation $n$ of the thermal noise in each cell.\\
$\muelF$       & Fourier transform matrix, of size $n_{\mathrm{pix}}\times n_{\mathrm{pix}}$.\\
$\muelS_b$     & Baseline selection matrix, which picks out 1\\
&  visibility out of the full set. Size $n_{\mathrm{pix}} \times n_b$\\
$\muelC_b$     & $n_{\mathrm{pix}} \times n_{\mathrm{pix}}$ convolution kernel that defines\\
               &  the PSF.\\
$\muelF_{bb'}$ & Convolution matrix mapping one $\deltu\deltv$ to $\deltl\deltm$.\\
& The set of all $\muelF_{bb'}$ determines the noise-PSF.\\
               & Size $n_{\mathrm{pix}} \times n_{\mathrm{pix}}$.\\
&\\
\hline
\end{tabular}
\caption{Table recapitulating the meaning and dimensions of vectors and matrices used {in Sec. \ref{sec.dudv}}. Only scalars which give matrix dimensions or indices are given here.}
\label{tab.variables}
\end{table}

@@

@@

\pg
This $\deltu\deltv$ space warrants further discussion: Fig. \ref{plot.psf.uv.dudv} shows, for a given $uv$-track (Fig. \ref{plot.psf.uv.dudv}b), both the corresponding $\deltu\deltv$ domain (Fig. \ref{plot.psf.uv.dudv}c) and point-spread function (Fig. \ref{plot.psf.uv.dudv}a). {The symmetric, negative $uv$-track is treated as a separate track, and thus ignored. This means that we do not fully constrain the noise-PSF (since the covariance matrix of the symmetric track is simply the Hermitian of the first), but we do not seek to constrain it in this section, but rather to show that our results hold.}. We can see that the $\delta u \delta v$-tracks are symmetrical about the origin. The $\delta u \delta v$ space corresponding to a given $uv$-track can thus be most concisely described as a ``filled $uv$-track", with its boundaries defined by the ends of the $uv$-track. The set of $\muelF_{bb'}$, each of which maps one value of the covariance matrix to a fringe in the image-plane, would then characterise a PSF equivalent for the noise distribution, which we refer to as the noise-PSF. In our formalism, the only source of statistical effects in the field are calibration errors and thermal noise. The average variance in all pixels will be given by the diagonal of the covariance matrix and the thermal noise, provided that it truly follows a normal distribution. The only effects which will cause the variance in the image-plane to vary from one pixel to the next are those mapped onto the covariance fringes, i.e. such position-dependent variance fluctuations will be caused by correlated gain errors, which are spurious signal introduced by erroneous gain estimates. Assuming all sources in the field are point-like and distant, then these variance fluctuations will follow a specific distribution, convolved to every source in the field.

\begin{figure*}[h!]
\centering
\includegraphics[width=\textwidth]{images/psfIM-dudv-labelled.png}
\caption{\label{plot.psf.uv.dudv} Fig. \ref{plot.psf.uv.dudv}a shows the PSF image of a simulated 1Jy source at phase centre. {Colourbar units are in Jansky}.  Fig. \ref{plot.psf.uv.dudv}b shows the associated UV track, and Fig. \ref{plot.psf.uv.dudv}c the corresponding $\left(\deltu,\deltv\right)$ tracks. Note that the $\deltu\deltv$ plane does not have a homogeneous point density, but is denser near its origin: here, this is shown by plotting only 1 random point in 10000.}
\end{figure*}

{Since the variance fluctuations act as tracers for calibration artefacts, artefacts in the image can be understood as \emph{one realisation} of the variance map, which is characterised by an average level determined by the variance in the gains and thermal noise, and a noise-PSF convolved with the sky brightness distribution}. The actual artefacts in the image will still be noisy, as a realisation of the true variance map. For the same reason, in the absence of correlated gain errors, $[\Cov{\matGains}]_{b \ne b'}$ are all zero and $\Cov{\matGains}$ is a diagonal matrix. {We then recover a ``flat" noise-map: the variance will be the same in all pixels, as the noise-PSF is absent}. In the ideal case, were we to recover the true value of the gains for all times and frequencies, this becomes pure thermal noise.
\subsection{Noise Map Simulations}\label{sec.simulations}

\begin{figure}[h!]
\centering
\includegraphics[width=0.8\textwidth]{images/CovarianceMatrix.png}
\caption{\label{fig.covmat} Example of {a non-stationary covariance matrix}, which can be used to simulate $\Cov{\matGains}$.{The colourbar units are Jy$^2$}. The correlation scale $\sigma_\tnuIndeix$ is 40 cells, and the variability period is 500 cells. The matrix is made positive semi-definite (and therefore a covariance matrix) through SVD decomposition. The maximum size of the ``bubbles" is determined by $\sigma_\tnuIndeix$.}
\end{figure}

@@

\begin{figure*}[h!]
\centering
\includegraphics[width=\textwidth]{images/cross_sections1.png}
\caption{\label{plot.cross-sections} The three lines in each figure correspond to three horizontal cross-sections from images in Fig. \ref{plot.noisepsfs}. {The units on the y-axis are dimensionless [$Jy^2/Jy^2$].} $\sigma_\tnuIndeix$ is the maximum characteristic error correlation length. In decreasing intensity, they correspond to  $m=0''$, $m=4''$, and $m=8''$. The dashed lines correspond to the variance measured with 2000 realisations for each pixel, while the solid line corresponds to the predicted value at that pixel. There are 31 pixels. We do not show cross-sections for negative $m$ due to image symmetry about the origin.}
\end{figure*}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{images/noisepsf-ctime400.png} %{../../../SkyCoverageMap.png} %
\caption{\label{plot.noisepsfs} Simulated noise-maps, compared with theoretical prediction. The pixel values are normalised by the average value of the covariance matrix: the units of the colourbar are thus dimensionless ($Jy^2/Jy^2$). These are on the same angular scale as the source shown in Fig. \ref{plot.psf.uv.dudv}.}
\end{figure}

@@

@@

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{images/Ctime6400-noisePSFandDirty-uncorr.png}
\caption{\label{imag.simu-3sources.noisemap} {Noise-map of sky with correlated gain errors and three point sources. The colourbars of (a), (b) and (c) have dimensionless units, while that of (d) is in Jansky. Note the presence of structure in the residuals (c): these show the limits of our hypothesis that sources are spatially incoherent.}}
\end{figure}

@@

\pg
We find that one $\bm{w}$ which satisfies the above is:
\begin{align}
\bm{w} &= \Cov{\matGains}^{-1} \mathbf{1}\label{eq.fullweights.analytical}
\end{align}
where $\mathbf{1}$ is a vector of ones. Provided that the hypotheses made so far hold\footnote{Specifically, that the residual visibilities do not contain any true signal, i.e. that all sources have been properly subtracted from the visibilities}, these weights depend only on calibration quality: badly-calibrated cells will include spurious time-correlated signal introduced by trying to fit the noise $n$ on visibilities. Down-weighting these cells helps suppress artefacts in the field, at the cost of far-field sensitivity. This weighting scheme is thus only a function of the relative quality of calibration at different times, boosting better-calibrated visibilities and suppressing poorly-calibrated visibilities. For the remainder of this paper, we will refer to these weights as \emph{artefact-optimal} weighting.

@@

\pg
We define the residual matrix as:
\begin{align}
\Rmat{\tnuIndeix} =& \sum_d s_d \Kmat{d,\tnuIndeix} \left( \ghatvect_\tnuIndeix\ghatvect_\tnuIndeix^H - \gvect_\tnuIndeix\gvect_\tnuIndeix^H \right) \Kmat{d,\tnuIndeix}^H + \epsilon \label{eq.residual.matrixform}
\end{align}
where we {explicitly place ourselves in the limits of our formalism, i.e.} that we do not have direction-dependent gains. We now see that at the core of Eq. \ref{eq.residual.matrixform} lies $\displaystyle\VarmatTau$, where $\sum_{\tnuIndeix\in\cellindex} \VarmatTau = \VarmatC$.
The K-matrix is defined as follows:
\begin{align}\label{kmat.def}
\Kmat{d,\tnuIndeix} =&
\begin{pmatrix}
k_{p,\tnuIndeix}^{d}   &                       &   0     &\\
                       & k_{q,\tnuIndeix}^{d}  &        &\\ 
         0              &                       & \ddots 
\end{pmatrix}
\end{align}
Since the residual matrix depends on the gains, we define the residual visibility vectors as:
\begin{align}
\Rmat{{\cellindex}} =&   \begin{pmatrix} \vdots \\ \Rmat{\tnuIndeix\in \cellindex} \\ \vdots \end{pmatrix} \\
\bm{R} =&   \begin{pmatrix} \hdots \quad \Rmat{c} \quad \hdots \end{pmatrix}
\end{align}

@@

@@

@@

\begin{figure}[h!]
    \centering
        \centering
        \includegraphics[width=\linewidth]{images/ctime001-NoisePsfCrossections-backup.png} 
        \caption{The sensitivity-optimal (green) and artefact-optimal (red) weights both give improvements over the unweighted noise-map (blue).} \label{fig.simu.weightapplied}
\end{figure}

@@

@@

\begin{figure}[h!]
\begin{subfigure}{.43\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/paperfig-goodcal.png}}
\caption{\label{image.3c295.goodcal-2} Well-calibrated, unweighted restored image of the sky near the centre of the Extended Groth Strip. Used for comparison with the other images. Calibration solution intervals used were 8 seconds, half the bandwidth. rms=5.87mJy/beam}
\end{subfigure}
\hfill
\begin{subfigure}{.43\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/paperfig-antweights.png}}
\caption{\label{image.3c295.antcorr} Image made using the same imaging parameters and corrected visibilities as Fig. \ref{image.3c295.nocorr}, with the application of antenna-based, sensitivity-optimal weighting. Solution interval of 2 minutes, half the bandwidth. rms=6.69mJy/beam}
\end{subfigure}
\caption{\label{image.twoRealImages}Comparison between the well-calibrated image (i.e. the same image as Fig \ref*{image.3c295.goodcal}) and antenna-based sensitivity-optimal weights.  {Units of both colourbars are Jansky}. We see that we recover a very similar image, despite the fact that the data used for the weighted image are averaged by a factor of 8 compared to those used for the unweighted image.}
\end{figure}
